---
title: "Reading dlf files"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Reading dlf files}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
# nolint start: indentation-linter
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
# nolint end
```

```{r setup}
library(daisyrVis)
```

There are are three functions for reading dlf files `read_dlf`, `read_dlf_dir` and `read_daisy_spawn_output`. `read_dlf` reads a single dlf file, `read_dlf_dir` searches a directory tree and reads all dlf files it finds, `read_daisy_spawn_output` reads the outputs from spawn program in `daisy`.

## Reading a single dlf file

We start by reading a single dlf file and inspect the contents. We use the Soil chemical log that is bundled with `daisyrVis` and read it with `read_dlf`.

```{r}
data_dir <- system.file("extdata", package="daisyrVis")
path <- file.path(data_dir, "hourly/P2D-Daily-Soil_Chemical_110cm.dlf")
dlf <- read_dlf(path)
```

`read_dlf` returns an S4 object of class `Dlf`. The data in the dlf file is stored in three _slots_ in the `Dlf` object, `header`, `units`, `data`
```{r}
slotNames(dlf)
```

The `header` slot is a list of the metadata that is stored before the actual data in the dlf file. The `units` slot is a `data.frame` with unit information for each data column. The `data` slot is a `data.frame` containing the data. A slot is accessed with the `@` operator
```{r}
dlf@header
```

```{r}
dlf@units
```

```{r}
head(dlf@data)
```

We can also directly access and assign columns of the `data` slot with the `$` and `[[` operators

```{r}
head(dlf$year)
dlf$year <- dlf$year + 2
head(dlf[["year"]])
dlf[[1, "year"]] <- dlf[[1, "year"]] + 10
dlf[[1, "year"]]
```

## Reading a directory with dlf files
We use the tracer and field nitrogen logs that are bundled with `daisyrVis` and read them with `read_dlf_dir`
```{r}
data_dir <- system.file("extdata", package="daisyrVis")
path <- file.path(data_dir, "annual")
```
Notice that the path is to a directory, instead of to a file as we did when reading a single dlf file.

```{r}
dlfs <- read_dlf_dir(path)
```
`read_dlf_dir` returns a list of `Dlf` objects
```{r}
names(dlfs)
```
To get a specific dlf we just index the list with the name
```{r}
name <- names(dlfs)[1]
name
dlfs[[name]]@header
```
The names are the file paths, which can be a bit unwieldly. We can either manually change the names or use the convenience functions `drop_dir_from_names` and `strip_common_prefix_from_names`
```{r}
dlfs <- drop_dir_from_names(dlfs)
names(dlfs)
```

```{r}
dlfs <- strip_common_prefix_from_names(dlfs)
names(dlfs)
```

```{r}
names(dlfs) <- c("A", "B", "C", "D", "E", "F", "G", "H")
names(dlfs)
dlfs[["A"]]@header
```

### Controlling which files are included
`read_dlf_dir` takes an optional pattern parameter that can be used to control which files are included. By default all files with the suffix `.dlf` are included. If we only want to read files ending with `2b.dlf`, we can do
```{r}
data_dir <- system.file("extdata", package="daisyrVis")
path <- file.path(data_dir, "annual")
pattern <- ".*2b\\.dlf"
dlfs <- read_dlf_dir(path, pattern)
names(dlfs)
```
`pattern` is a regular expression. You should around with it and see if you can find a shorter pattern that results in the same files being read.


## Reading Daisy spawn output
Daisy spawn produces a directory hierarchy where each directory contains the same log files. We use the Daisy spawn outputs that are bundled with `daisyrVis` and read them with `read_daisy_spawn_output`
```{r}
data_dir <- system.file("extdata", package="daisyrVis")
path <- file.path(data_dir, "daisy-spawn-like")
list.files(path, recursive=TRUE)
```
Notice that each directory contains the same log files.
```{r}
dlfs <- read_daisy_spawn_output(path)
```
As `read_daisy_dir`, `read_daisy_spawn_output` also returns a list of `Dlf` objects
```{r}
dlf_names <- names(dlfs)
dlf_names
```
But notice that instead of a list with an entry for each file that was read, we have a list with an entry for each log type that was read. All log of the same type are concatenated together, and an extra column is added to the data field. The name of the column is "sim" by default, but can be set by passing `col_name="MySimulationName"` to `read_daisy_spawn_output`
```{r}
name <- dlf_names[1]
head(dlfs[[name]]@data)
unique(dlfs[[name]]$sim)
```
