---
title: "Plotting dlfs"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Plotting dlfs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
# nolint start: indentation-linter
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
  )
# nolint end
```

```{r setup}
library(daisyrVis)
```

There is a general plot function `plot_dlf` as well as some convenience wrappers around `plot_dlf` that simplifies some basic tasks. Here we focus on `plot_dlf`.

To get started we will load some dlf files with `read_dlf`
```{r}
data_dir <- system.file("extdata", package="daisyrVis")
dlfs <- read_dlf(file.path(data_dir, "annual/Annual-FN"))
dlfs <- strip_common_prefix_from_names(dlfs)
```
These dlfs contain annualy logged variables related to field nitrogen
```{r}
colnames(dlfs[[1]]@data)
```

We can plot all the variables, or a subset. Here we plot "Surface_Loss" and "Denitrification".
```{r}
y_vars <- c("Surface_Loss", "Denitrification")
plot_dlf(dlfs, "year", y_vars, type="bar")
```
Notice how `plot_dlf` plots each variable in a separate sub plot, and each scenario (AKA dlf) with a different color.


We can also make a line or points plot, or combine the plots types
```{r}
plot_dlf(dlfs, "year", y_vars, "lines")
plot_dlf(dlfs, "year", y_vars, "points")
plot_dlf(dlfs, "year", y_vars, "lines-points-bar")
```

The lines and points plot types are usually better suited when we have more data points. To see this we will load another dlf file using `read_dlf`
```{r}
path <- file.path(data_dir, "hourly/P2D-Daily-Soil_Chemical_110cm.dlf")
dlf <- read_dlf(path)
colnames(dlf@data)
```
```{r}
y_vars <- c("Leak_Matrix", "Transform")
plot_dlf(dlf, "time", y_vars, "points")
```

If we only want to plot a part of the time period we can use `subset_dlf`
```{r}
summer95 <- subset_dlf(dlf, "1995-06-01", "1995-08-31")
plot_dlf(summer95, "time", y_vars, "lines", title_suffix=" - Summer of '95")
```
