---
title: "Reading dlf files"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{reading-dlf-files}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
# nolint start: indentation-linter
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
# nolint end
options(rmarkdown.html_vignette.check_title = FALSE)
```

```{r setup}
library(daisyrVis)
```

There are are two functions for reading dlf files `read_dlf` and `read_dlf_dir`. `read_dlf` reads a single dlf file, `read_dlf_dir` searches a directory tree and reads all dlf files it finds. `read_dlf_dir` is particularly useful for reading outputs the spawn program in `daisy`.

## Reading a single dlf file

We start by reading a single dlf file and inspect the contents. We use the Soil chemical log that is bundled with `daisyrVis` and read it with `read_dlf`.

```{r}
data_dir <- system.file("extdata", package="daisyrVis")
path <- file.path(data_dir, "hourly/P2D-Daily-Soil_Chemical_110cm.dlf")
dlf <- read_dlf(path)
```

`read_dlf` returns an S4 object of class `Dlf`. The data in the dlf file is stored in three _slots_ in the `Dlf` object, `header`, `units`, `body`
```{r}
slotNames(dlf)
```

The `header` slot is a list of the metadata that is stored before the actual data in the dlf file. The `units` slot is a `data.frame` with unit information for each data column. The `body` slot is a `data.frame` containing the data. A slot is accessed with the `@` operator
```{r}
dlf@header
```

```{r}
dlf@units
```

```{r}
head(dlf@body)
```

## Reading a directory with dlf files
We use the tracer and field nitrogen logs that are bundled with `daisyrVis` and read them with `read_dlf_dir`
```{r}
data_dir <- system.file("extdata", package="daisyrVis")
path <- file.path(data_dir, "annual")
```
Notice that the path is to a directory, instead of to a file as we did when reading a single dlf file.

```{r}
dlfs <- read_dlf_dir(path)
```
`read_dlf_dir` returns a list of `Dlf` objects
```{r}
names(dlfs)
```
To get a specific dlf we just index the list with the name
```{r}
name <- names(dlfs)[1]
name
dlfs[[name]]@header
```
The names are the file paths, which can be a bit unwieldly. We can either manually change the names or use the convenience functions `drop_dir_from_names` and `strip_common_prefix_from_names`
```{r}
dlfs <- drop_dir_from_names(dlfs)
names(dlfs)
```

```{r}
dlfs <- strip_common_prefix_from_names(dlfs)
names(dlfs)
```

```{r}
names(dlfs) <- c("A", "B", "C", "D", "E", "F", "G", "H")
names(dlfs)
dlfs[["A"]]@header
```

### Controlling which files are included
`read_dlf_dir` takes an optional pattern parameter that can be used to control which files are included. By default all files with the suffix `.dlf` are included. If we only want to read files ending with `2b.dlf`, we can do
```{r}
data_dir <- system.file("extdata", package="daisyrVis")
path <- file.path(data_dir, "annual")
pattern <- ".*2b\\.dlf"
dlfs <- read_dlf_dir(path, pattern)
names(dlfs)
```
`pattern` is a regular expression. You should around with it and see if you can find a shorter pattern that results in the same files being read.
